---
title: blog test post
author: Alexander C. Hungerford
date: '2020-11-29'
slug: []
categories:
  - blog
tags:
  - fantasy football
  - beginners
description: ''
featured: ''
featuredalt: ''
featuredpath: ''
linktitle: ''
type: post
---
## Setup

```{r setup}
# packages to load
library(tidyverse)
```

## Our Fantasy Football League

### Basics

* League name: HFFFL
* 12 participants: referred to as GMs (general managers)
* Draft: Auction style, $100
* Dynasty League: our rosters don't reset each year. Each GM keeps the same number of players on his roster every year, after the 2019 season that was 3 *keepers* per team.
* Keepers: 3 per team (high, mid, low). High keepers were players drafted for \$15+. Mid keepers were drafted for \$5-\$15. Low keepers were drafted for \$4 or less. You can categorize keepers up, but not down. So a GM could retain 3 "low" players from one season to the next, but each would be assigned a designation of high, mid, or low.

## League Basics


```{r}
# league members - not sorted
gm <- c("Alex", "Katz", "Brett", "Mikulich", "Chris", "Jack", "Jon", "Todd", "Ian", "Ashton", "Dave", "Nate")

# alphabetized gm list
gm <- sort(gm)

# show list
gm
```

I would like to create a table of data that shows who placed 1st, 2nd, and 3rd each year.

```{r}
# list of years league has been active
year <- c(2012:2019)

# podium placement, by year
first <- c("Mikulich", "Nate", "Nate", "Chris", "Nate", "Nate", "Dave", "Dave")
second <- c("Jon", "Alex", "Ian", "Todd", "Jon", "Chris", "Jon", "Jon")
third <- c("Katz", "Katz", "Jack", "Dave", "Katz", "Katz", "Chris", "Todd")

# podium tibble
podium <- tibble(year, first, second, third)

# show work
podium
```

I'd like to see the distribution of Top 3 finishes by each GM. `count()` will show me the distribution for a given variable (first, second, third), but only in separate outputs.

```{r}
# distribution of first place finishes
podium %>% count(first)
```

```{r}
# distribution of second place finishes
podium %>% count(second)
```

I can create a new tibble from `podium`, but the problem is that not every GM is listed because two (Ashton & Brett) have never placed in the Top 3.


```{r}
podium_counts <-
   podium %>%
   pivot_longer(cols = -year) %>%
   count(name, value) %>%
   pivot_wider(names_from = name, values_from = n, values_fill = list(n = 0))

podium_counts
```

I think I have to create another tibble to show all the GMs and count their Top 3 finishes.

```{r}

```

